How to select predictive models for decision-making or causal inference?

### 2023-12-16 how_to_select_causal_models.tex
iter

    \includegraphics[width=1\linewidth]{images/toy_random_forest_high_R2_high_tau_risk.pdf}%

@@

The Neyman-Rubin Potential Outcomes framework
\cite{naimi2023defining,imbens_causal_2015} enables statistical reasoning on
causal treatment effects: Given an outcome $Y \in \mathbb R$ (\eg mortality risk
or hospitalization length), function of a binary treatment $A \in \mathcal{A} =
    \{0, 1\}$ (\eg a medical procedure), and baseline
covariates $X \in \mathcal{X} \subset \mathbb{R}^d$, we observe the factual
distribution, $O = (Y(A), X, A) \sim \mathcal D = \mathbb P(y, x, a)$. However,
we want to model the existence of potential observations (unobserved ie.
counterfactual) that correspond to a different treatment. Thus we want
quantities on the counterfactual distribution $O^{*} = (Y(1), Y(0), X, A) \sim
    \mathcal D^{*} = \mathbb P(y(1), y(0), x, a)$.

>>>>

    \setlength{\unitlength}{.1\linewidth}
    \begin{picture}(10,5.2)(0,0)
    \put(0, 0){\includegraphics[width=1\linewidth]{images/toy_random_forest_high_R2_high_tau_risk.pdf}}%
    \put(7.05,4.4){\vector(0,1){.47}}
    \put(7.05,4.4){\vector(0,-1){.55}}
    \put(7.1,4.5){\scriptsize\sffamily treatment effect}
    \put(7.1,4.15){\small $\tau(x)$}
    \put(5.2,3.6){\small $\hat{\tau}(x)$}
    \put(5.15,3.6){\vector(0,1){.29}}
    \put(5.15,3.6){\vector(0,-1){.15}}
    \end{picture}

@@

The Neyman-Rubin Potential Outcomes framework
\cite{naimi2023defining,imbens_causal_2015} enables statistical reasoning on
causal treatment effects: Given an outcome $Y \in \mathbb R$ (\eg mortality risk
or hospitalization length), function of a binary treatment $A \in \mathcal{A} =
    \{0, 1\}$ (\eg~a medical procedure), and baseline
covariates $X \in \mathcal{X} \subset \mathbb{R}^d$, we observe the factual
distribution, $O = (Y(A), X, A) \sim \mathcal D = \mathbb P(y, x, a)$. However,
we want to model the existence of potential observations (unobserved ie.~counterfactual) that correspond to a different treatment. Thus we want
quantities on the counterfactual distribution $O^{*} = (Y(1), Y(0), X, A) \sim
    \mathcal D^{*} = \mathbb P(y(1), y(0), x, a)$.
